Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies along a climatic gradient in the rear-edge

### 2019-07-10 man/ecosystems/ms_ecosystems.Rmd
update

We found similar tree competence levels in our sampled sites (Table 1), but differences in tree size and age suggest different management origin (*i.e.* land-use legacy). The review of historical documents revealed that our study sites had different land-use trajectories driven mainly by the differential pattern of the natural resource use [@JimenezOlivencia2015; @MartinCivantos2014]. On the one hand, the northern sites (*e.g.* San Juan), show a sequentially distribution of the land uses along the elevational gradient [@JimenezOlivencia2015]: grasslands and shrublands for cattle farming are located at high elevations; then forests formation with some croplands; and at lowlands, irrigated terraces with tree crops. Other activities (*e.g.* mining) could also determine the use of natural resources, and therefore the forest structure. For instance, the SJ oak woodland is located in an area with a high concentration of mines and quarries that have been exploited intermittently throughout history. Historical documents indicated two periods of intense mining activity: the second half of the 19th century after the publication of detailed mineralogical reports by [@Maestre1858]; and during the first decades of the twentieth century until 1960, which is the last year in which there is evidence of the existence of mining activity. On the other hand, woodland areas of the southern slopes (Cáñar) are mixed with a greater percentage of croplands, even reached high elevation (mainly barley, rye and potatoes) [@Calatrava2019]. Forest resources, like firewood, charcoal and acorns, have been continuously exploited in southern sites through history [@Valbuena2013; @JimenezSerrano2004]. In Cáñar site, there were an arraigated tradition of charcoal extraction since fifteenth century [@JimenezSerrano2004] up to middle of the 20th century. From this moment to the present, there were a sharp decrease of the wood extraction, mainly due to rural abandonment [@MesaTorres2009; @Bonet2014_conama]. Acorn resources have been tradiotionally exploited in this location up to 1950s [@MesaTorres2009; @Bonet2014_conama].

>>>>

We found similar tree competence levels in our sampled sites (Table 1), but differences in tree size and age suggest different management origin (*i.e.* land-use legacy). The review of historical documents revealed that our study sites had different land-use trajectories driven mainly by the differential pattern of the natural resource use [@JimenezOlivencia2015; @MartinCivantos2014]. On the one hand, the northern sites (*e.g.* San Juan), show a sequentially distribution of the land uses along the elevational gradient [@JimenezOlivencia2015]: grasslands and shrublands for cattle farming are located at high elevations; then forests formation with some croplands; and at lowlands, irrigated terraces with tree crops. Other activities (*e.g.* mining) could also determine the use of natural resources, and therefore the forest structure. For instance, the SJ oak woodland is located in an area with a high concentration of mines and quarries that have been exploited intermitently throughout history. Historical documents indicated two periods of intense mining activity: the second half of the 19th century after the publication of detailed mineralogical reports by [@Maestre1858]; and during the first decades of the twentieth century until 1960, which is the last year in which there is evidence of the existence of mining activity. On the other hand, woodland areas of the southern slopes (Cáñar) are mixed with a greater percentage of croplands, even reached high elevation (mainly barley, rye and potatoes) [@Calatrava2019]. Forest resources, like firewood, charcoal and acorns, have been continuously exploited in southern sites through history [@Valbuena2013; @JimenezSerrano2004]. In Cáñar site, there were an arraigated tradition of charcoal extraction since fifteenth century [@JimenezSerrano2004] up to middle of the 20th century. From this moment to the present, there were a sharp decrease of the wood extraction, mainly due to rural abandonment [@MesaTorres2009; @Bonet2014_conama]. Acorn resources have been tradiotionally exploited in this location up to 1950s [@MesaTorres2009; @Bonet2014_conama].
